---
title: ''
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysis of data sets

## 1. 

## (a) 

There are 3 lines contain bunch of "1"s and some comments in the data set. I delete these 3 lines because i think they are meaningless.

\bigskip

## (b)

```{r}
cmort = read.table("cmort.dat")
plot(cmort$V1, col = 'red', type = 'l', 
     xlab = 'time', ylab = 'Average cardiovascular mortality rate',
     main = 'Average cardiovascular mortality rate in Los Angeles California \n versus time')
```

\bigskip

## (c)

### i trend
There is a slightly downward trend. The average cardiovascular mortality rate goes up a lot when time is around 150.

### ii periodicity and seasonality 
The data seems periodic but it does not appear to be seasonality.

### iii heteroskedasticity
The variability seems slightly decreasing.

### iv dependence 
There might be some negative dependency.

### v outliers, missing data, etc.
There is no obvious outliers, and missing data.



\newpage

## 2.

## (a)

There is nothing strange in this data file, so i did not make any adjustment. 

\bigskip

## (b)

```{r}
varve = read.table('varve.dat')
plot(varve$V1, col = 'red', type = 'l',
     xlab = 'time', ylab = 'Glacial varve thickness in Maine',
     main = 'Glacial varve thickness versus time')
```

\bigskip

## (c)

### i trend
There is no obvious trend. The Glacial varve thickness goes up extremely when time is around 80. This might be a outlier.

### ii periodicity and seasonality 
There is no obvious periodicity and seasonality.

### iii heteroskedasticity
The variablity seems increasing after time is aournd 350.

### iv dependence 
There might be some negative dependency.

### v outliers, missing data, etc.
data point around time is 80 might be a outlier which need further investigation.


\newpage

## 3.

## (a)

There are 4 dates in the end of data set, and keep them will affect how R read our data, so i delete these 4 dates.

\bigskip

## (b)

```{r}
choc = read.table('choc.dat')
par(mar = c(5.1,5.1,4.1,2.1))
plot(choc$V1, col = 'red', type = 'l',
     xlab = 'time', 
     ylab = 'Monthly chocolated-based confectionery production \n (in tones)',
     main = 'Australian monthly chocolate-based confectionery production \n versus time')
```

\bigskip

## (c)

### i trend
There is a slightly upward trend with structural break aournd time is 150 and 250.

### ii periodicity and seasonality 
There is a seasonality on a 12 month interval.

### iii heteroskedasticity
The variability seems increasing.

### iv dependence 
There is an oscillatory pattern. There might be a negative dependency.

### v outliers, missing data, etc.
There is no obvious outliers, and missing data.